---
output: github_document
---

# Cycle Hire Locations in London


This repo contains reproducible code to get London cycle hire data.

We will get them using R with the following packages loaded:

```{r, message=FALSE, warning=FALSE}

library(tidyverse)
library(osmdata)
library(tmap)
tmap_mode("view")

```


## Official data

TfL's official dataset on cycle hire locations is no longer available: https://data.london.gov.uk/dataset/tfl-cycle-hire-locations

However, there was a FOI made in 2020, leading to the publication of cycle hire location data, which can be accessed from R as follows:

```{r}
u = "http://foi.tfl.gov.uk/FOI-2898-1920/FOI-2898-1920.xlsx"
f = basename(u)
download.file(u, f)
stations_tfl_2020 = readxl::read_excel(f)
stations_tfl_2020
nrow(stations_tfl_2020)
```
As the above code shows, that contains 802 docking stations.
We believe these are stations that are currently active.

We set the id to a character to match the OSM data:

```{r}
stations_tfl_2020$official_ref = as.character(stations_tfl_2020$Station.Id)
```

We will remove the pop-up stations, which only exist for a few days, as follows:

```{r}
stations_tfl_2020 = stations_tfl_2020 %>% 
  filter(!stringr::str_detect(string = StationName, pattern = "Pop"))
```



## OSM data

The OSM community has been active in mapping many things, including cycle hire docking stations.
The following commands get this data:

```{r}
stations_osm_list = opq("london") %>% 
  add_osm_feature(key = "amenity", value = "bicycle_rental") %>% 
  osmdata_sf()

stations_osm = stations_osm_list$osm_points 
stations_osm = stations_osm %>% select(osm_id, name, brand, capacity, start_date, network, official_name, official_ref)


stations_osm = stations_osm %>% 
  filter(str_detect(string = network, "tfl") | str_detect(string = brand, "antand"))

```

## Comparing TfL and OSM sources

There are 8 more stations in the official dataset. 
We can join the OSM data with the official data as follows, after a few checks:

```{r}
summary(duplicated(stations_osm$official_ref))
summary({osm_in_official = stations_osm$official_ref %in% stations_tfl_2020$official_ref})
nrow(stations_tfl_2020) - nrow(stations_osm)
stations_combined = left_join(
  stations_tfl_2020,
  stations_osm %>% sf::st_drop_geometry()
  )
```

<!-- We can also add a dataset approximating the first wave of the data as follows: -->

```{r, echo=FALSE}
# stations_2010 = spData::cycle_hire_osm
# stations_2010$year = 2010
# stations_combined = left_join(
#   stations_combined,
#   stations_2010 %>% sf::st_drop_geometry() %>% select(name, year)
# )
```


```{r}
stations_combined$capacity = as.numeric(stations_combined$capacity)
```

We need to make this combined dataset a geographic class for mapping:

```{r}
# stations_sf = sf::st_as_sf(stations_combined, coords = c("longitude", "latitude"), crs = 4326)
stations_sf = sf::st_as_sf(stations_combined, coords = c("Easting", "Northing"), crs = 27700)
```


```{r plot}
stations_area = stations_sf %>% 
  sf::st_buffer(1000) %>% 
  sf::st_union()

qtm(stations_area, fill = NULL) +
  tm_shape(stations_sf) +
  tm_dots(size = "capacity", scale = 0.3, alpha = 0.5)
```

## Updating the data

There are many ways we could improve the data.
The most obvious is the start date.
If anyone has ideas on that, please let us know on the issue tracker.

## Publishing open data

We hope to collaborate with others to improve and eventually publish data data as open data.



